Isle of Wight Employment Stats
========================================================

This is an example of pulling down jobless figures from nomisweb for the Isle of Wight.

Provided as a demonstrator for OnTheWight by Tony Hirst, (Computing and Communications Department, The Open University, and the Open Knowledge Foundation).

```{r echo=FALSE, message=FALSE}
#Package loading
list.of.packages <- c("XML", "RCurl", "ggplot2", "xtable")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(XML)
library(RCurl)
library(ggplot2)
library(xtable)
```
```{r echo=FALSE, message=FALSE}
#Utility functions
commaReplace = function(node) {
  val = xmlValue(node)
  ans = gsub(",", "", val)
  if(is.na(ans))
    val
  else
    ans
}

comma=function(txt){paste(txt,',',sep='')}
```
```{r echo=FALSE, message=FALSE}
#Total stats loader
#total1, male1, female1
tt=postForm("https://www.nomisweb.co.uk/reports/lmp/la/2038431803/subreports/jsa_time_series/report.aspx", .encoding='utf-8', pivot="total1")
tables <- readHTMLTable(tt,stringsAsFactors = FALSE,elFun=commaReplace ) #encoding = "UTF-8")
nn=gsub('([\n\t\r])','',names(tables$`NULL`))
names(tables$`NULL`)=nn
ud=tables$`NULL`
ud['Isle of Wight']=lapply(ud['Isle of Wight'], function(x) as.integer(x))
ud['South East(%)']=lapply(ud['South East(%)'], function(x) as.numeric(x))
ud['Great Britain(%)']=lapply(ud['Great Britain(%)'], function(x) as.numeric(x))
#We need to create a date. IF we just the 1/MON/YEAR the chart labeling maybe looks better than the possibly more correct lastDayOfMonth/MON/YEAR? Or use middle of month (15th, say?)?
ud['tDate']=lapply(ud['Date'], function(x) as.Date(paste(28,x),"%d %b %Y"))
```

The most recent figures:

```{r results='asis', message=FALSE, echo=FALSE}
print(xtable(ud[nrow(ud),c(1,2,4,5)]),include.rownames=FALSE,type='html')
```

The previous figures:

```{r results='asis', message=FALSE, echo=FALSE}
print(xtable(ud[nrow(ud)-1,c(1,2,4,5)]),include.rownames=FALSE,type='html')
```

Last three months' figures:
```{r results='asis', message=FALSE, echo=FALSE}
df.tmp=ud[(nrow(ud)-2):nrow(ud),c(1,2,4,5)]
print(xtable(df.tmp),include.rownames=FALSE,type='html')
```

A little bit of machine generated explanatory text:

```{r results='asis', message=FALSE, echo=FALSE, comment='' }
txt="The total number of people claiming Job Seeker's Allowance (JSA) on the Isle of Wight in"
txt=paste(txt,format(ud[nrow(ud),]$tDate,'%B'),'was',ud[nrow(ud),'Isle of Wight'])
txt=comma(txt)

jsa.lm=ud[nrow(ud)-1,'Isle of Wight']
jsa.diff.m=ud[nrow(ud),'Isle of Wight']-jsa.lm
if (jsa.diff.m>0){
  txt=paste(txt,'up',jsa.diff.m,'from',jsa.lm)
} else if (jsa.diff.m<0) {
  txt=paste(txt,'down',abs(jsa.diff.m),'from',jsa.lm)
} else { txt=paste(txt,'the same as') }
txt=paste(txt,'in',format(ud[nrow(ud)-1,]$tDate,'%B, %Y,'))

txt=paste(txt,'and')
jsa.ly=ud[nrow(ud)-12,'Isle of Wight']
jsa.diff.y=ud[nrow(ud),'Isle of Wight']-jsa.ly
if (jsa.diff.y>0){
  txt=paste(txt,'up',jsa.diff.y,'from',jsa.ly)
} else if (jsa.diff.y<0) {
  txt=paste(txt,'down',abs(jsa.diff.y),'from',jsa.ly)
} else { txt=paste(txt,'the same as') }
txt=paste(txt,'in',format(ud[nrow(ud)-12,]$tDate,'%B, %Y.'))

cat(txt)
```


```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
ud$dy <- unlist(format(ud["tDate"],'%Y'))
ud$dm <- unlist(format(ud["tDate"],'%b'))
ud$dm=factor(ud$dm)
ud$dm<- factor(ud$dm, levels = month.abb)

ggplot(ud)+geom_line(aes_string(x='dy',y='`Isle of Wight`',group='dm'))+facet_grid(~dm)+theme(axis.text.x = element_text(angle = 90, hjust = 1))
#reorder(variable, value)
```

```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
ggplot(ud[(nrow(ud)-11):nrow(ud),])+geom_line(aes_string(x='tDate',y='`Isle of Wight`'))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
```