Isle of Wight Employment Stats
========================================================

This is an example of pulling down jobless figures from nomisweb for the Isle of Wight.

Provided as a demonstrator for OnTheWight by Tony Hirst, (Computing and Communications Department, The Open University, and the Open Knowledge Foundation).

```{r echo=FALSE, message=FALSE}
#Package loading
list.of.packages <- c("XML", "RCurl", "ggplot2", "xtable")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(XML)
library(RCurl)
library(ggplot2)
library(xtable)
```
```{r echo=FALSE, message=FALSE}
#Utility functions
commaReplace = function(node) {
  val = xmlValue(node)
  ans = gsub(",", "", val)
  if(is.na(ans))
    val
  else
    ans
}

comma=function(txt){paste(txt,',',sep='')}
```
```{r echo=FALSE, message=FALSE}
#Total stats loader
#total1, male1, female1
tt=postForm("https://www.nomisweb.co.uk/reports/lmp/la/2038431803/subreports/jsa_time_series/report.aspx", .encoding='utf-8', pivot="total1")
tables <- readHTMLTable(tt,stringsAsFactors = FALSE,elFun=commaReplace ) #encoding = "UTF-8")
nn=gsub('([\n\t\r])','',names(tables$`NULL`))
names(tables$`NULL`)=nn
ud=tables$`NULL`
ud['Isle of Wight']=lapply(ud['Isle of Wight'], function(x) as.integer(x))
ud['South East(%)']=lapply(ud['South East(%)'], function(x) as.numeric(x))
ud['Great Britain(%)']=lapply(ud['Great Britain(%)'], function(x) as.numeric(x))
#We need to create a date. IF we just the 1/MON/YEAR the chart labeling maybe looks better than the possibly more correct lastDayOfMonth/MON/YEAR? Or use middle of month (15th, say?)?
ud['tDate']=lapply(ud['Date'], function(x) as.Date(paste(28,x),"%d %b %Y"))
```

The most recent figures:

```{r results='asis', message=FALSE, echo=FALSE}
print(xtable(ud[nrow(ud),c(1,2,4,5)]),include.rownames=FALSE,type='html')
```

The previous figures:

```{r results='asis', message=FALSE, echo=FALSE}
print(xtable(ud[nrow(ud)-1,c(1,2,4,5)]),include.rownames=FALSE,type='html')
```

Last three months' figures:
```{r results='asis', message=FALSE, echo=FALSE}
df.tmp=ud[(nrow(ud)-2):nrow(ud),c(1,2,4,5)]
print(xtable(df.tmp),include.rownames=FALSE,type='html')
```

Previous year's figures:
```{r results='asis', message=FALSE, echo=FALSE}
df.tmp=ud[(nrow(ud)-14):(nrow(ud)-11),c(1,2,4,5)]
print(xtable(df.tmp),include.rownames=FALSE,type='html')
```

A little bit of machine generated explanatory text:

```{r results='asis', message=FALSE, echo=FALSE, comment='' }
txt="The total number of people claiming Job Seeker's Allowance (JSA) on the Isle of Wight in"
txt=paste(txt,format(ud[nrow(ud),]$tDate,'%B'),'was',ud[nrow(ud),'Isle of Wight'])
txt=comma(txt)

jsa.lm=ud[nrow(ud)-1,'Isle of Wight']
jsa.diff.m=ud[nrow(ud),'Isle of Wight']-jsa.lm
if (jsa.diff.m>0){
  txt=paste(txt,'up',jsa.diff.m,'from',jsa.lm)
} else if (jsa.diff.m<0) {
  txt=paste(txt,'down',abs(jsa.diff.m),'from',jsa.lm)
} else { txt=paste(txt,'the same as') }
txt=paste(txt,'in',format(ud[nrow(ud)-1,]$tDate,'%B, %Y,'))

txt=paste(txt,'and')
last.yr.row=nrow(ud)-12
jsa.ly=ud[last.yr.row,'Isle of Wight']
jsa.diff.y=ud[nrow(ud),'Isle of Wight']-jsa.ly
if (jsa.diff.y>0){
  txt=paste(txt,'up',jsa.diff.y,'from',jsa.ly)
} else if (jsa.diff.y<0) {
  txt=paste(txt,'down',abs(jsa.diff.y),'from',jsa.ly)
} else { txt=paste(txt,'the same as') }
txt=paste(txt,'in',format(ud[last.yr.row,]$tDate,'%B, %Y.'))

cat(txt)
```


```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
ud$dy <- unlist(format(ud["tDate"],'%Y'))
ud$dm <- unlist(format(ud["tDate"],'%b'))
ud$dm=factor(ud$dm)
ud$dm<- factor(ud$dm, levels = month.abb)

g=ggplot(ud[ud$dy>2010,])+geom_line(aes_string(x='dm',y='`Isle of Wight`',group='dy',col='dy'))+theme(axis.text.x = element_text(angle = 90, hjust = 1))
g=g+ylab('Isle of Wight JSA claimants')
g
```


```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
library(directlabels)
g=ggplot(ud[ud$dy>2010,],aes_string(x='dm',y='`Isle of Wight`'))+geom_line(aes_string(group='dy',col='dy'))+theme(axis.text.x = element_text(angle = 90, hjust = 1))+scale_x_discrete(expand=c(0, 1.5))
#direct.label(g)
#In the following, new to increase internal viewport so labels don't overflow right
g=g+ylab('Isle of Wight JSA claimants')
direct.label(g,'last.points')
```

If we plot the monthly change we get a better idea for how monthly changes compare with previous years:

```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
ud$l=c(NA, diff(ud$`Isle of Wight`, lag = 1))
g=ggplot(ud[ud$dy>2010,])+geom_line(aes_string(x='dm',y='l',group='dy',col='dy'))+theme(axis.text.x = element_text(angle = 90, hjust = 1))
g=g+ylab('Isle of Wight JSA claimants - monthly change')
print(g)
```

```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
ggplot(ud)+geom_line(aes_string(x='dy',y='`Isle of Wight`',group='dm'))+facet_grid(~dm)+theme(axis.text.x = element_text(angle = 90, hjust = 1))
#reorder(variable, value)
```

```{r fig.width=7, fig.height=6, message=FALSE, echo=FALSE}
ggplot(ud[(nrow(ud)-11):nrow(ud),])+geom_line(aes_string(x='tDate',y='`Isle of Wight`'))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
nomis <- read.csv('http://www.nomisweb.co.uk/api/v01/dataset/NM_1_1.data.csv?geography=1946157281&date=latestMINUS1,latest&select=date_name,geography_name,geography_code,sex_name,item_name,measures_name,obs_value,obs_status_name&measures=20100,20203')


nomis <- read.csv('http://www.nomisweb.co.uk/api/v01/dataset/NM_4_1.data.csv?geography=1946157281&date=latestMINUS1,latest&measures=20100&select=DATE,DATE_NAME,GEOGRAPHY_NAME,AGE_DUR_NAME,SEX_NAME,MEASURES,MEASURES_NAME,OBS_VALUE')
nx=nomis[grep('Claiming',nomis$AGE_DUR_NAME),]
nx[SEX_NAME=='Total,']

nx$AGE_DUR_NAME <- factor(nx$AGE_DUR_NAME, levels = c("Claiming one week or less", "Claiming over 1 and up to 2 weeks", "Claiming over 2 and up to 4 weeks" , "Claiming over 4 and up to 6 weeks" , "Claiming over 6 and up to 8 weeks" ,"Claiming over 8 and up to 13 weeks", "Claiming over 13 and up to 26 weeks","Claiming over 26 and up to 39 weeks","Claiming over 39 and up to 52 weeks", "Claiming over 52 and up to 65 weeks","Claiming over 65 and up to 78 weeks" ,"Claiming over 78 and up to 104 weeks" ,"Claiming over 104 and up to 156 weeks" , "Claiming over 156 and up to 208 weeks","Claiming over 208 and up to 260 weeks" ,"Claiming over 260 weeks" ))

nxxx=cast(nxx[,c('DATE','AGE_DUR_NAME','OBS_VALUE')],AGE_DUR_NAME~DATE)


```
