The Rwandan Tragedy
========================================================

*A reworking of a data investigation produced by Alfred Essa/@malpaso [The Rwandan Tragedy: Data Analysis with 7 Lines of Simple Python Code](http://alfredessa.com/2013/05/the-rwandan-tragedy-data-analysis-with-7-lines-of-simple-python-code/)*


The first step is to get some life expectancy [data from the World Bank](http://datacatalog.worldbank.org/) - it seems like there's a World Bank Data service API wrapper for that:

```{r}
#install.packages("WDI")
library(WDI)
```

Let's see if we can find a code for life expectancy?
```{r}
WDIsearch(string='life.*expectancy', field='name', cache=NULL)
```

Ah ha, seems like "SP.DYN.LE00.IN" (Life expectancy at birth, total (years)) will do it...

(There are also codes for life expectancy for males and females separately)

```{r}
df.le=WDI(country="all", indicator=c("SP.DYN.LE00.IN"), start=1900, end=2012)

```

We'll be doing some charting, so let's use ggplot... Load the required library...
```{r  message=F}
require(ggplot2)
```

Alfred used a boxplot to provide an overview of the range of life expectanices across countries over a period of year. The outliers during the 1990s really jumped out:

```{r fig.width=7, fig.height=6, message=F, warning=FALSE} 
g=ggplot()+geom_boxplot(data=df.le,aes(x=year,y=SP.DYN.LE00.IN, group=year)) 
g=g+theme(axis.text.x = element_text(angle = 45, hjust = 1))
g
``` 

Let's filter the data to tunnel down and look to see which country or countries  the outliers correspond to:

```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
subset(df.le,year>1988 & SP.DYN.LE00.IN<40)
```

Rwanda is notable, so let's overlay the numbers for life expectancy in Rwanda on the chart:

```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
g=g+geom_line(data=subset(df.le,country=='Rwanda'),aes(x=year,y=SP.DYN.LE00.IN),col='red')
g
``` 

So what's causing the drop life expectancy? One way of exploring this problem is to look at the life expectancy figures for other countries with known problems over a particular period to see if their life expectancy figures have a similar signature over that particular period.

So for example, let's bring in in data for Kenyan life expectancy - does the Aids epidemic that hit that country have a similar signarture effect?

```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
g=g+geom_line(data=subset(df.le,country=='Kenya'),aes(x=year,y=SP.DYN.LE00.IN),col='green')
g
``` 
How about Uganda, which suffered similarly?

```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
g=g+geom_line(data=subset(df.le,country=='Uganda'),aes(x=year,y=SP.DYN.LE00.IN),col='blue')
g
``` 
Neither of those traces appear to have the same signature as the Rwandan curve. So might there be another cause? How about civil war? For example, Bangladesh suffered a civil war in the early 1970s - what was the effect on life expectancy over that period?
```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
g=g+geom_line(data=subset(df.le,country=='Bangladesh'),aes(x=year,y=SP.DYN.LE00.IN),col='purple')
g
```
Ah ha - that has a marked similarity, to the eye at least...

Search for mortaility indicators
```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
WDIsearch(string='mortality', field='name', cache=NULL)
```
Grab "Mortality rate, under-5 (per 1,000 live births)" overall, and broken down for male and female
```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
df.cm=WDI(country="all", indicator=c("SH.DYN.MORT","SH.DYN.MORT.FE","SH.DYN.MORT.MA"), start=1900, end=2009)
``` 
Overall boxplot
```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
gx=ggplot(df.cm)+geom_boxplot(aes(x=year,y=SH.DYN.MORT, group=year))
gx
```
Does Rwanda account for any of the outliers? Let's overlay the Rwanda stats
```{r fig.width=7, fig.height=6, message=F, warning=FALSE}
gx=gx+geom_line(data=subset(df.cm,country=='Rwanda'),aes(x=year,y=SH.DYN.MORT),col='red')
gx
``` 


Hmmm.. some but not all... So what are the outliers in this case? Do you think you could find out?